Analyzing Personal Movement by Users of Activity Monitoring Devices
===================================================================


```{r, results="hide", echo=TRUE}

library(plyr)
library(chron)
library(ggplot2)
require(lattice)
library(gridExtra)

df <- read.csv("activity.csv")

df$date <-as.Date(df$date,"%Y-%m-%d")

by_day <- ddply(df, .(format(date, "%m-%d")), summarize, total=sum(steps, na.rm=TRUE), mean=mean(steps, na.rm=TRUE))

colnames(by_day) <- c("Date", "Total", "Mean")

by_day <- na.omit(by_day)

```

# Histogram of total steps taken each day

```{r fig.width=7, fig.height=6,  echo=TRUE}

hist(by_day$Total, xlab="Steps", main="Histogram of total steps taken each day")

```

## Mean and median of total steps by day

```{r,  echo=TRUE}

mean(by_day$Total)

median(by_day$Total)

```

## Average daily activity pattern?

```{r, each=TRUE}

by_five_mins <- ddply(df, .(interval), summarize, mean=mean(steps, na.rm=TRUE))

qplot(interval, mean, data=by_five_mins, geom="line", xlab="5-mins Interval", ylab="Mean Steps", main="Average number of steps across 5-min intervals")

```

## Maximum steps interval

```{r, echo=TRUE}

rw <- which.max(by_five_mins$mean)

by_five_mins$interval[rw]

```

## Adjust for NA values

```{r, echo=TRUE}

# How many NAs are there?

navalues <- apply(df, 2, function(x) length(which(is.na(x))))

```

###There are `r navalues[1]` NA rows in the original activity data.

###Use mean for the day as number for missing data

```{r, echo=TRUE}

df2 <- df

replace_mean <- function(x) replace(x, is.na(x), mean(x, na.rm = TRUE))

df3 <- ddply(df2, ~ interval, transform, steps = replace_mean(steps))

```

## Histogram of total steps taken each day with missing data filled in

```{r fig.width=7, fig.height=6,  echo=TRUE}

by_day_filled <- ddply(df3, .(format(date, "%m-%d")), summarize, total=sum(steps, na.rm=TRUE), mean=mean(steps, na.rm=TRUE))

colnames(by_day_filled) <- c("Date", "Total", "Mean")

hist(by_day_filled$Total, xlab="Steps", main="Histogram of total steps taken each day")

```

## Mean and median of total steps by day with missing data filled in

```{r,  echo=TRUE}

mean(by_day_filled$Total)

median(by_day_filled$Total)

```

```{r fig.width=7, fig.height=6, echo=TRUE}

df3$day <- weekdays(as.Date(df3$date))

five_min_weekend <- df3[which(df3$day=="Saturday" | df3$day=="Sunday"),]

by_five_min_both <- ddply(five_min_weekend, .(interval), summarize, weekend=mean(steps, na.rm=TRUE))


five_min_weekday <- df3[which(df3$day!="Saturday" & df3$day!="Sunday"),]

by_five_min_weekday <- ddply(five_min_weekday, .(interval), summarize, mean=mean(steps, na.rm=TRUE))

by_five_min_both$weekday <- by_five_min_weekday$mean

plot1 <- xyplot(weekday ~ interval, data=by_five_min_both, type="l")
plot2 <- xyplot(weekend ~ interval, data=by_five_min_both, type="l")
grid.arrange(plot1,plot2, nrow=2)

```
